Performing graph-based clustering with rOpenScPCA

### 2024-12-16 analyses/hello-clusters/01_perform-evaluate-clustering.Rmd
simpler saving

# Calculate clusters with Seurat
seurat_obj <- seurat_obj |>
  SCTransform() |>
  RunPCA() |>
  FindNeighbors() |>
  FindClusters()

@@

Results can either be directly exported as a TSV file (e.g., with `readr::write_tsv()`), or you can add the results into your SCE or Seurat object.
The subsequent examples will show you how to save the `cluster_results_df` data frame to an SCE and a Seurat object.

_A word of caution!_
Objects from the ScPCA Portal already contain a column called `cluster` with results from an automated clustering.
These automatic clusters were not evaluated, and their parameters were not optimized for any given library.

To avoid ambiguity between the existing and new clustering results, you may want to rename incoming columns.

We'll therefore begin by adding the prefix `ropenscpca_` to all column names in the `cluster_results_df` data frame before adding them to a given object.

```{r rename columns}
# First, rename columns in `cluster_results_df` to avoid ambiguity
cluster_results_renamed_df <- cluster_results_df |>
  # add the prefix "ropenscpca_" to all columns
  dplyr::rename_with(\(x) paste0("ropenscpca_", x))

cluster_results_renamed_df
```

@@

Follow these steps to save the clustering results to an SCE object's `colData`:

```{r add to sce}
# First, extract the colData into a separate data frame for manipulation
sce_coldata <- colData(sce) |>
  as.data.frame()

# Second, add columns as desired using the barcodes and cell_id columns to join
sce_coldata <- sce_coldata |>
  dplyr::left_join(
    cluster_results_renamed_df,
    by = c("barcodes" = "ropenscpca_cell_id")
  )

# Third, add the modified colData back into the SCE
# Use the row.names argument to ensure rownames (barcodes) are retained; don't forget this step!
colData(sce) <- DataFrame(
  sce_coldata,
  row.names = sce_coldata$barcodes
)

head(colData(sce))
```

@@

Follow these steps save the clustering results to a Seurat object's cell metadata.
Note that the Seurat function `AddMetaData()`, used below, assumes that the row order in the provided data frame matches the order of cells in the Seurat object you're adding metadata to.

```{r add to seurat}
# Add the cluster result data frame to the Seurat object metadata
seurat_obj <- AddMetaData(seurat_obj, cluster_results_renamed_df)

# Print the cell metadata
head(seurat_obj@meta.data)
```

>>>>

# Calculate clusters with Seurat using a standard Seurat pipeline, for example
seurat_obj <- seurat_obj |>
  SCTransform() |>
  RunPCA() |>
  FindNeighbors() |>
  FindClusters()

@@

Results can either be directly exported as a TSV file (e.g., with `readr::write_tsv()`), or you can add the results into your SCE or Seurat object.
The subsequent examples will demonstrate saving the cluster assignments stored in `cluster_results_df$cluster` to an SCE and a Seurat object.

_A word of caution!_
Objects from the ScPCA Portal already contain a column called `cluster` with results from an automated clustering.
These automatic clusters were not evaluated, and their parameters were not optimized for any given library.
To avoid ambiguity between the existing and new clustering results, we'll name the new column `ropenscpca_cluster`.

@@

We can add columns to an SCE object's `colData` table by directly creating a column in the object with `$`.
Before we do so, we'll confirm that the clusters are in the same order as the SCE object by comparing cell ids:

```{r check sce order}
all.equal(
  colnames(sce),
  cluster_results_df$cell_id
)
```

```{r add to sce}
# Add cluster results to the colData
sce$ropenscpca_cluster <- cluster_results_df$cluster
```

@@

We can add columns to an Seurat object's cell metadata table by directly creating a column in the object with `$` (note that you can also use the Seurat function `AddMetaData()`).
Before we do so, we'll confirm that the clusters are in the same order as the Seurat object by comparing cell ids:

```{r check seurat order}
all.equal(
  colnames(seurat_obj),
  cluster_results_df$cell_id
)
```

```{r add to seurat}
# Add cluster results to the cell metadata
seurat_obj$ropenscpca_cluster <- cluster_results_df$cluster
```
